# Introduction {#intro}


**bookdown** est un package R développé par Yihui Xe qui permet d’écrire des livres en format Rmarkdown pour les publier en PDF ou HTML. Le principe de base consiste à utiliser un fichier Rmarkdown par chapitre.


**bookdown** peut être installé depuis le CRAN 

```{r cran, eval=FALSE}
install.packages("bookdown")
```

ou depuis le repo GitHub

```{r remotes, eval=FALSE}
# install.packages("remotes")
remotes::install_github("rstudio/bookdown")
```


Une fois ceci fait, on peut créer un fichoer via `File > New project > New directory > Book Project using bookdown` et choisir l'emplacement et le nom du futur livre.

Un exemple minimal de livre est ainsi crée à cet emplacement et peut être compilé en cliquant sur ` Build > Build Book`

On peut alors s'inspirer des fichiers déjà créés.









You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].
