---
title: "Uilisation GitHub"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

Git is a version management software. The idea is to save the different changements made on a projet, and create an historic of it. To do so, between two saved versions, changements will be added on git with the "add" command, and a little comment on the modification will be joint with the "commit" command as explained later.

Github is a remote service, to save the different git versions. It also enable several people to work on a same project. To use it, you have to :   

* create a Git hub account
* make a link (ssh) between your computer and your account
* create or join a repository
* use the "push" and "pull" commands to share your work and get the advanced from others. 


The goal of this little course is to get the basis of this tool, and its use. 


## 1. Get started
### 1.1 Installation of git 
#### On Linux and MacOS :
To use git on Linux or MacOS, you'll need to install the software with some command like :

```{bash}
$ sudo apt-get update
$ sudo apt-get install git
```


### 1.2 GitHub :
#### Create an account
First of all, you need to create a github account

#### Create a link between your account and your computer with a SSH Key :


#### Where is my project ? 
###### On one of my repository
###### Join an existing project


## 2. Basis commands 

### Add a file in your repository 
Add a local document named "utilisation_GitHub.Rmd" in your git repository. You can display the differences between the local file and the current commit with git status.
```{bash}
$ git add utilisation_GitHub.Rmd
$ git status
```

### Commit a file in your repository 
Create a new commit containing the current contents of the index and the given log message describing the changes. The following command is used to record what has been changed in the document :
```{bash}
$ git commit -m "Add my latest version of the document"
```
The modified file is not on the server yet. 

### Synchronizing the file in the server
Upload your file to the server:
```{bash}
$ git push
```
To upload the files from the server to your local repository.
```{bash}
$ git pull
```


The different steps can be summarized by the figure displayed at the beginning.




## 3. Confilit 

```{r}

```

